\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{float}
\usepackage{natbib}
\bibliographystyle{apsr}
\usepackage{graphicx}
\graphicspath{ {../fig/} }
\usepackage{setspace}
\setstretch{2}

\begin{document}

\title{Revise and Resubmit Memo}
\author{Robert Lytle}
\date{\today}
\maketitle

\noindent The principle shortcomings of my initial draft were its brevity and its vaguery. To reconcile this, I make the following changes:
\begin{enumerate}
\item I correct a misrepresentation of Levendusky's \textit{The Partisan Sort} in Section 2.
\item I add a single paragraph to the end of Section 2, briefly summarizing the aims of Iyengar, Sood, and Lelkes. Where appropriate, I make an attempt to succinctly reiterate the goals of the authors more frequently throughout the piece, as they are many.
\item I clarify the opening paragraph of Section 3.2, expanding on ``Where Iyengar, Sood and Lelkes (2012) excels in conceptual clarity, it falls short in theoretical coherence." The meaning of this sentence was previously fairly arcane.
\item I recreate the DAG in black and white to address accessibility concerns.
\item I stress that a shortcoming of the paper being replicated is it's unclear articulation of the data generating process.
\item In Section 3.3.1 I specify the \textit{measures} used to operationalize partisan affect and generally reorganize these paragraphs to be more coherent.
\item I expanded section 3.3 significantly, describing a natural experiment using a Geographic Regression Discontinuity design. Previously, the section was overly-pessimistic towards the prospect of gaining causal leverage over media.
\item In section 3.3.5 I recognize SUTVA as problematic to the authors' causal inference.
\item I correct ambiguous language about the regression model used by the authors in section 3.3.6.
\item I expand my proposed subgroup analysis significantly. I clarify my hypotheses and theoretical claims and present additional data intended to motivate and justify additional study.
\end{enumerate}
In addition to these largely substantive changes, I correct grammatical and typographical errors throughout the piece. I am thankful to all those who provided helpful feedback and constructive criticism---I believe this draft to be a significant improvement over my initial submission.


\thispagestyle{empty}
\clearpage
%\setstretch{1}
%\bibliography{references}

\end{document}

%